Digital Analysis of Paintings

### 2014-05-01 paper.tex
Add paper 9 details.

@@

>>>>

Often these techniques involve using a form of machine learning to gather
meaningful data from the high dimensional feature space which image processing
processes provide.

@@

All artists have a distinguishing style which may change as their career
progresses. Art historians find this a challenging problem as many factors of
the painting need to be taken into account. Image processing and machine
learning techniques can be used to aid this process.

Stylometry, the study of an artists style, has many problems involved with it.
Two of these are: extracting distinguishing features and dating a
painting, both these problems are tackled in \cite{jafarpour2009stylistic}.

This research uses considers the artists style to be a hidden variable which
controls the observable properties of the image; the colour, brushstrokes, etc.

The dating challenge is the act of dating a painting from a catalogue of known
paintings from the artists lifetime. An artists style is likely to change over
time, especially as they meet their peers and try out new movements.

This research considers how a human expert might approach the problem of
determining style and that paintings may degrade over time due to the materials
used to create them. Historians must combine a potentially noisy observation
with pre-existing knowledge to come up with an analysis of the style.

Computer systems have to use features of an image, considering both global and
local information, to build up a high-dimensional feature space which can then
have statistical analysis performed upon it to gauge some information about
style.

It has already been discussed that a \gls{HSL} colour space provides a more
representative view of an image than a \gls{RGB} space. However, \gls{HSL} does
not exist in a Cartesian space. This can be achieved by transforming the
\gls{HSL} space into an XYZ space by ``unrolling'' the radial elements of
\gls{HSL}.

As with many techniques which utilise wavelet transforms, a very high
resolution is used to gather information which may be too fine for a human eye
to perceive. A dual-tree complex wavelet transform using the aforementioned XYZ
colour space can detect colour patterns as well as local difference.

These wavelets provide coefficients in very high dimensional space which a
large amount of noise. To deal with this there is a need for dimensionality
reduction and normalisation of noise within the feature space.

Hidden Markov Trees provide an image model over numerous different resolutions.
They act like Hidden Markov Models but also considers a tree like structure
which maps to dyadic space. This allows for an image to be described as a
statistical entity rather than a set of pixels.

At each scale there are hidden variables which control the wavelet
coefficients. These hidden variable represent either a smooth region, which has
a small variance, or an edge, which has a large variance.

The images are split into several patches to allow the analysis to be performed
effectively.

Results for the dating challenge, 66 Van Gogh paintings were used with the goal
of dating three test images which cannot be easily dated by art historians.
Training was performed on the set of paintings using 10-fold cross-validation
using several different classifiers. The best classifier; random forest; had a
generalisation performance of 73.7\% and the results from the three test images
aligned with the conclusions of the art historians, but this is somewhat
conjecture as the dates aren't officially known.

Extraction distinguishing features focused on extracting flowers within the
paintings. Irrelevant patches were removed from the training images and again
training was performed using 10-fold cross-validation using several
classifiers. Once again random forest was the best decision and several
distinct distinguishing features.

The results of this research do seem promising, and the techniques described
here do seem like they can provide some very powerful analysis.

The sample size is a little small, but the split into different patches allows
for a larger number of samples for training. Though this still only has a
certain number of date ranges to work with.

The size of the patches is never fully explored and the research blindly uses
$256 \times 256$ pixels for each region. It would be interesting to see what
effect changing the size of these regions would make upon the research.

In conclusion this research provides some very powerful analysis with the
experimental results to back it up in two different fields of stylometry which
agree with expert opinion and begin to bring forth some new ideas.
